Supporting Information to Estimating the Initial Susceptible Fraction in Dengue Dynamics"

### 2014-11-26 S1_lm1.tex
started writing about the stochastic multi-strain model in S1

This approach thus makes possible to incorporate epidemiological knowledge 
about disease biology (e.g. the magnitude of $R_0$) into the computation of \rr.
This may prove particularly important when disease counts are low and/or close 
to the detection threshold.
\newpage
\bibliography{lm1}
\end{document}

>>>>

This approach thus makes possible to incorporate epidemiological knowledge 
about disease biology (e.g. the magnitude of $R_0$) into the computation of \rr.
This may prove particularly important when disease counts are low and/or close 
to the detection threshold.
\newpage
\section*{Multi-strain dynamics}

In order to validate the estimates derived in the paper, we simulated 
multi-strain Dengue dynamics by means of a stochastic 4-serotype SIR model with 
cross-immunity as described below. The model permits up to 4 dengue infections 
with reduced susceptibility after the first Dengue episode due to cross 
immunity. Immunity to each serotype is considered complete and permanent. 
Figure xx depicted all possible states and state-transitions included in the 
model.

Let $S$ be individuals susceptible to all 4 types of dengue, $I_i$ infectious 
with Dengue type $i$ and $R_i$ individuals recovered from Dengue type $i$. 
Infectious individuals already on their secondary and later Dengue infections 
are represented by multiple indices. For example, $I_{[23]1}$ is an individual 
which has had Dengues type 2 and 3 in the past -- and therefore is immune to 
them -- and is currently transmitting Dengue 1. The index outside the bracket 
denotes current infection. Recovered individuals indices denote their immunity, 
so for instance $R_{123}$ is an individual which is immune to Dengue types 1, 2 
and 3, but not to 4. Let $I_{*i} = \sum I_{[\ldots]i}$ with $[\ldots]$ 
representing exposure history of the infected individual which can vary from 0 
to 3 in length. All individuals are born to the $S$ state and birth and death 
rates are equal.

The possible state-transitions and their  propensities are listed in table 
\ref{tab:trans}.
\begin{table}
\caption{
\bf{State-transitions and propensities}}
\label{tab:trans}
\begin{center}
\begin{tabular}[c]{l|l|l}
\hline
Transition & Propensity & Description\\
\hline
$S \rightarrow I_i$ & $\beta S I_{*i}$ & Primary infection \\
$I_i \rightarrow R_i$ & $\sigma I_i$ & Primary recovery\\
$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & Secondary infection\\
$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ & Secondary recovery\\
$R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ & Tertiary 
infection\\
$I_{[ij]k} \rightarrow R_{ijk}$ & $\sigma I_{[ij]k}$ & Tertiary recovery\\
$R_{ijk} \rightarrow I_{[ijk]l}$ & $\beta \delta R_{ijk} I_{*l}$ & Quaternary 
infection\\
$I_{[ijk]l} \rightarrow R_{ijkl}$ & $\sigma I_{[ijk]l}$ & Quaternary recovery\\
$\rightarrow S$ & $\mu N$ & Birth\\
$All \rightarrow$ & $\mu N$ & Death\\
\hline
\end{tabular}
\end{center}

\end{table}

\bibliography{lm1}
\end{document}
